### 2022-03-04 paper/GHRVis.tex
draft: latex structure change

\documentclass{egpubl}
\usepackage{eurovis2021}

% --- for  Annual CONFERENCE
% \ConferenceSubmission   % uncomment for Conference submission
% \ConferencePaper        % uncomment for (final) Conference Paper
% \STAR                   % uncomment for STAR contribution
% \Tutorial               % uncomment for Tutorial contribution
% \ShortPresentation      % uncomment for (final) Short Conference Presentation
% \Areas                  % uncomment for Areas contribution
% \MedicalPrize           % uncomment for Medical Prize contribution
% \Education              % uncomment for Education contribution
% \Poster                 % uncomment for Poster contribution
% \DC                     % uncomment for Doctoral Consortium
%
% --- for  CGF Journal
% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
% \JournalPaper         % uncomment for final version of Journal Paper
%
% --- for  CGF Journal: special issue
% \SpecialIssueSubmission    % uncomment for submission to , special issue
\SpecialIssuePaper         % uncomment for final version of Computer Graphics Forum, special issue
%                          % EuroVis, SGP, Rendering, PG
% --- for  EG Workshop Proceedings
% \WsSubmission      % uncomment for submission to EG Workshop
% \WsPaper           % uncomment for final version of EG Workshop contribution
% \WsSubmissionJoint % for joint events, for example ICAT-EGVE
% \WsPaperJoint      % for joint events, for example ICAT-EGVE
% \Expressive        % for SBIM, CAe, NPAR
% \DigitalHeritagePaper
% \PaperL2P          % for events EG only asks for License to Publish

% --- for EuroVis 
% for full papers use \SpecialIssuePaper
% \STAREurovis   % for EuroVis additional material 
% \EuroVisPoster % for EuroVis additional material 
% \EuroVisShort  % for EuroVis additional material

% !! *please* don't change anything above
% !! unless you REALLY know what you are doing
% ------------------------------------------------------------------------

\usepackage[T1]{fontenc}
\usepackage{dfadobe}

% \usepackage{cite}  % comment out for biblatex with backend=biber
% ---------------------------
\biberVersion
\BibtexOrBiblatex
\usepackage[backend=biber,bibstyle=EG,citestyle=alphabetic,backref=true,url=false]{biblatex}
%\addbibresource{egbibsample.bib}

\renewbibmacro*{doi+eprint+url}{%
    \printfield{doi}%
    \newunit\newblock%
    \iftoggle{bbx:eprint}{%
        \usebibmacro{eprint}%
    }{}%
    \newunit\newblock%
    \iffieldundef{doi}{%
        \usebibmacro{url+urldate}}%
        {}%
    }

\addbibresource{GeoVis.bib}

\newcommand{\citetitlea}[1]{\citefield{#1}{title}}

% ---------------------------  
\electronicVersion
\PrintedOrElectronic
% for including postscript figures
% mind: package option 'draft' will replace PS figure by a filename within a frame
\ifpdf \usepackage[pdftex]{graphicx} \pdfcompresslevel=9
\else \usepackage[dvips]{graphicx} \fi

\usepackage{egweblnk}
% end of prologue

\usepackage{float,hhline,hyperref,array,tabulary,tabularx,etoolbox,multirow}

\usepackage{dblfloatfix}

\usepackage[table, dvipsnames]{xcolor}

\usepackage{multicol}

\include{preamble/preamble.basic}
\include{preamble/preamble.main}
\include{preamble/preamble.math}
\include{preamble/preamble.draft}

\author[Q. Wang \& R. S. Laramee]
{\parbox{\textwidth}{\centering Q. Wang
  and R. S. Laramee
    }
    \\
\parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK}
}

\title{Cartograms with Dynamic Features}

\begin{document}

\pagestyle{plain}

\maketitle

\begin{abstract}

\end{abstract}

\input{section/introduction}

\input{section/related}

\input{section/data}

\input{section/main}

\section{Evaluation}

\section{Limitations and Future Work}

\section{Conclusions}

\section{Acknowledgments}
This work is funded by the grant EP/S010238/2 from the Engineering and Physical Sciences Research Council (EPSRC). EPSRC is a British Research Council that provides government funding for grants to undertake research and postgraduate degrees in engineering and the physical sciences.

% reset \section command for section title "References"
\let\section=\origsection
\printbibliography

\end{document}