### 2022-11-09 paper/section/conclusion.tex

\section{Conclusions}

In this paper, we present a novel approach to generate cartograms with rivers. We first propose a new algorithm to generate cartograms with rivers, and then we present a prototype to support the exploration of cartograms with rivers. We evaluate our approach using a user study, and the results partially support our hypotheses: the introduction of rivers improves the legibility but doesn't reduce the time needed to locate a region. We also discuss the limitations of our work and the future directions.