### 2022-11-17 paper/section/conclusion.tex
draft: review

\section{Conclusions}

In this paper, we present a novel approach to generate non-contiguous cartograms with rivers. We first propose a new algorithm to generate cartograms with rivers, and then we present a prototype to support the exploration of cartograms with rivers. We evaluate our approach with a preliminary user study, and the results support our hypotheses: the introduction of rivers improves the legibility and recognizability of cartograms, although a deeper analysis is necessary to reach more confident conclusions. We also discuss the limitations of our work and future research directions.