### 2023-07-06 paper/GHRVis.tex
draft: IV

\documentclass[Afour,sagev,times]{sagej}

\usepackage{moreverb,url}

\usepackage{float,hhline,array,tabulary,tabularx,etoolbox,multirow, multicol}

\newcommand{\algorithmautorefname}{Algorithm}

\usepackage[table, dvipsnames]{xcolor}

\include{preamble/preamble.main}
\include{preamble/preamble.math}

\usepackage[colorlinks,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage[nameinlink]{cleveref}

\setcounter{secnumdepth}{3}

\begin{document}

\title{Demers Cartogram with Rivers}

% \includegraphics[width=\textwidth,keepaspectratio]{figure/cover.png}
%         \captionof{figure}{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}

\begin{abstract}
    Cartograms serve as representations of geographical and abstract data, employing a value-by-area mapping technique. A relative of the Dorling cartogram, the Demers cartogram, utilizing squares instead of circles to represent regions. This alternative approach allows for a more intuitive comparison of regions, utilizing screen space more efficiently. However, a drawback of the Dorling cartogram and its variants lies in the potential displacement of regions from their original positions, ultimately compromising legibility, readability, and accuracy. To tackle this limitation, we propose a novel hybrid cartogram layout algorithm that incorporates dynamic topological elements, such as rivers, into Demers cartograms. The presence of rivers significantly impacts both the layout and visual appearance of the cartograms. Through a user study conducted on an Electronic Health Records (EHR) dataset, we evaluate the efficacy of the proposed hybrid layout algorithm. The obtained results illustrate that this approach successfully retains key aspects of the original cartogram while enhancing legibility, readability, and overall accuracy.
\end{abstract}

\keywords{Cartogram, Demers Cartogram}

\maketitle

\input{section/introduction}

\input{section/related}

\input{section/data}

\input{section/main}

\input{section/evaluation}

\input{section/limitation}

\input{section/conclusion}

% \bibliographystyle{mslapa}
\urlstyle{same}
\bibliographystyle{plainnat}
\bibliography{references}

\cleardoublepage

\input{section/appendix}

\end{document}